
##### ADT DATA #####

```{r}
library(Seurat)
library(Matrix)
library(SeuratDisk)
library(SeuratObject)
```

```{r}
# Setting work directory so that reading data is relative to this notebook
wd <- setwd(".")
```

```{r}
# Reading in ADT object written to file by analysis notebook
ADT_seurat <- readRDS(paste(wd, "ADT_seurat_analysis_output/seurat_objects", "seurat.rds", sep="/"))
```

### ADDING CELL TYPE ANNOTATION FROM REFERENCE DATASET

```{r}
#Get list of barcodes needed for subsetting later
ADT_barcodes <- colnames(ADT_seurat)
```

```{r}
#Load in reference seurat object containing cell type annotation data
ref_srat <- LoadH5Seurat(paste(wd, "data/pbmc_multimodal.h5seurat", sep="/"))
```

```{r}
#extract cell type annotation data into a dataframe
celltype_df <- as.data.frame(ref_srat@meta.data[,"celltype.l1"])
```

```{r}
#rename rows to be the cell barcodes and rename column to be appropriat
rownames(celltype_df) <- rownames(ref_srat@meta.data)
colnames(celltype_df) <- "celltype"
```

```{r}
#filter celltype dataframe to only contain cells remaining from preprocessing in seurat object as reference seurat was not preprocessed
ADT_filt_celltype_df <- subset(celltype_df, rownames(celltype_df) %in% ADT_barcodes)
```

```{r}
#add cell type annotations to seurat
ADT_seurat@meta.data$celltype <- ADT_filt_celltype_df$celltype
```

```{r}
#plotting embedding with cell type annotation labels
DimPlot(ADT_seurat, reduction = "umap", group.by = "celltype", label = TRUE) + NoLegend()
```

### Integrating MCL clustering data into Seurat

```{r}
#Read MCL clustering table
ADT_mcl_table <- read.table(paste(wd, "ADT_mcl_work/mcl_seurat_table/ADT_srt_mcl.file", sep = "/"))
```

```{r}
#calculate length of table
ADT_mcl_len <- dim(ADT_mcl_table)[1]
```

```{r}
#create dataframe containing only clusters and remove dummy line that MCL needs
ADT_mcl_df <- as.data.frame(ADT_mcl_table$V2[2:ADT_mcl_len]) 
```

```{r}
#rename rows to be the cell barcodes and rename column to be appropriate
rownames(ADT_mcl_df) <- ADT_mcl_table$V1[2:ADT_mcl_len]
colnames(ADT_mcl_df) <- "MCL_clustering"
```

```{r}
#integrate mcl clustering to seurat object
ADT_seurat@meta.data$MCL_clustering <- ADT_mcl_df$MCL_clustering
```

```{r}
#plot full seurat object coloured by MCL clusters, legend has been removed as it is just arbitrary cluster numbers
DimPlot(ADT_seurat, reduction = "umap", group.by = "MCL_clustering") + NoLegend()
```

```{r}
#plot full seurat object coloured by Leiden clustersm legend has been removed as it is just arbitrary cluster numbers
DimPlot(ADT_seurat, reduction = "umap", group.by = "seurat_clusters") + NoLegend()
```

### Testing subsetting seurat by MCL clusterings > specified number of nodes

```{r}
# Counts how many cells are in each MCL cluster
ADT_cls_count <- as.data.frame(dplyr::count(ADT_seurat@meta.data, MCL_clustering))
```

```{r}
# Filter out clusters with less than specified number of cells within
ADT_cls_filt <- ADT_cls_count[ADT_cls_count$n >= 150,]
```

```{r}
#filter seurat object to only have cells in clusters with >= 150 cells
ADT_filt_srat <- subset(x = ADT_seurat, cells = rownames(subset(ADT_seurat@meta.data, MCL_clustering %in% ADT_cls_filt$MCL_clustering)))
```

```{r}
#plot filtered seurat object coloured by MCL clusters, legend has been removed as it is just arbitrary cluster numbers
DimPlot(ADT_filt_srat, reduction = "umap", group.by = "MCL_clustering") + NoLegend()
```

```{r}
#plot filtered seurat object coloured by leiden clusters, legend has been removed as it is just arbitrary cluster numbers
DimPlot(ADT_filt_srat, reduction = "umap", group.by = "seurat_clusters") + NoLegend()
```

#### TESTING VOLATILITY PLOTTING ####

```{r}
#Read volatility table
ADT_vol_table <- read.table(paste(wd, "ADT_mcl_work/comparison/vol.srt.file", sep = "/"))
```

```{r}
#calculate length of table
ADT_vol_len <- dim(ADT_vol_table)[1]
```

```{r}
#create dataframe containing only volatility scores and remove dummy line that MCL needs
ADT_vol_df <- as.data.frame(ADT_vol_table$V3[2:ADT_vol_len])
```

```{r}
#rename rows to be the cell barcodes and rename column to be appropriate
rownames(ADT_vol_df) <- ADT_vol_table$V2[2:ADT_vol_len]
colnames(ADT_vol_df) <- "Volatility"
```

```{r}
#Integrate volatility scores into seurat object
ADT_seurat@meta.data$Volatility <- ADT_vol_df$Volatility
```

```{r}
#plot volatility where a lower score (red) represents a cell more likely to change clusters between algorithms
FeaturePlot(ADT_seurat, features = "Volatility", label = FALSE, cols = c("red", "blue"))
```

```{r}
# Saving the seurat object so it can easily be read in for future analysis
dir.create(file.path(wd, "ADT_seurat_analysis_output", "integrated_seurat"), recursive = TRUE)
saveRDS(ADT_seurat, paste(wd, "ADT_seurat_analysis_output/integrated_seurat", "full_seurat.rds", sep="/"))
```

####### RNA DATA ########

```{r}
# Setting work directory so that reading data is relative to this notebook
wd <- setwd(".")
```

```{r}
# Reading in ADT object written to file by analysis notebook
RNA_seurat <- readRDS(paste(wd, "RNA_seurat_analysis_output/seurat_objects", "seurat.rds", sep="/"))
```

### ADDING CELL TYPE ANNOTATION FROM REFERENCE DATASET

```{r}
#Get list of barcodes needed for subsetting later
RNA_barcodes <- colnames(RNA_seurat)
```

```{r}
#filter celltype dataframe to only contain cells remaining from preprocessing in seurat object as reference seurat was not preprocessed
RNA_filt_celltype_df <- subset(celltype_df, rownames(celltype_df) %in% RNA_barcodes)
```

```{r}
#add cell type annotations to seurat
RNA_seurat@meta.data$celltype <- RNA_filt_celltype_df$celltype
```

```{r}
#plotting embedding with cell type annotation labels
DimPlot(RNA_seurat, reduction = "umap", group.by = "celltype", label = TRUE) + NoLegend()
```

### Integrating MCL clustering data into Seurat

```{r}
#Read MCL clustering table
RNA_mcl_table <- read.table(paste(wd, "RNA_mcl_work/mcl_seurat_table/RNA_srt_mcl.file", sep = "/"))
```

```{r}
#calculate length of table
RNA_mcl_len <- dim(RNA_mcl_table)[1]
```

```{r}
#create dataframe containing only clusters and remove dummy line that MCL needs
RNA_mcl_df <- as.data.frame(RNA_mcl_table$V2[2:RNA_mcl_len]) 
```

```{r}
#rename rows to be the cell barcodes and rename column to be appropriate
rownames(RNA_mcl_df) <- RNA_mcl_table$V1[2:RNA_mcl_len]
colnames(RNA_mcl_df) <- "MCL_clustering"
```

```{r}
#integrate mcl clustering to seurat object
RNA_seurat@meta.data$MCL_clustering <- RNA_mcl_df$MCL_clustering
```

```{r}
#plot full seurat object coloured by MCL clusters, legend has been removed as it is just arbitrary cluster numbers
DimPlot(RNA_seurat, reduction = "umap", group.by = "MCL_clustering") + NoLegend()
```

```{r}
#plot full seurat object coloured by Leiden clustersm legend has been removed as it is just arbitrary cluster numbers
DimPlot(RNA_seurat, reduction = "umap", group.by = "seurat_clusters") + NoLegend()
```

### Testing subsetting seurat by MCL clusterings > specified number of nodes

```{r}
# Counts how many cells are in each MCL cluster
RNA_cls_count <- as.data.frame(dplyr::count(RNA_seurat@meta.data, MCL_clustering))
```

```{r}
# Filter out clusters with less than specified number of cells within
## NOTE: 50 cells in RNA has similar number of clusters to 150 cells in ADT, meaning that ADT produced more finer clusters, 100 cells were used to ensure that some smaller clusters were lost but also not too many to make visualisation meaningless
RNA_cls_filt <- RNA_cls_count[RNA_cls_count$n >= 100,]
```

```{r}
#filter seurat object to only have cells in clusters with >= 100 cells
RNA_filt_srat <- subset(x = RNA_seurat, cells = rownames(subset(RNA_seurat@meta.data, MCL_clustering %in% RNA_cls_filt$MCL_clustering)))
```

```{r}
#plot filtered seurat object coloured by MCL clusters, legend has been removed as it is just arbitrary cluster numbers
DimPlot(RNA_filt_srat, reduction = "umap", group.by = "MCL_clustering") + NoLegend()
```

```{r}
#plot filtered seurat object coloured by leiden clusters, legend has been removed as it is just arbitrary cluster numbers
DimPlot(RNA_filt_srat, reduction = "umap", group.by = "seurat_clusters") + NoLegend()
```

#### TESTING VOLATILITY PLOTTING ####

```{r}
#Read volatility table
RNA_vol_table <- read.table(paste(wd, "RNA_mcl_work/comparison/vol.srt.file", sep = "/"))
```

```{r}
#calculate length of table
RNA_vol_len <- dim(RNA_vol_table)[1]
```

```{r}
#create dataframe containing only volatility scores and remove dummy line that MCL needs
RNA_vol_df <- as.data.frame(RNA_vol_table$V3[2:RNA_vol_len])
```

```{r}
#rename rows to be the cell barcodes and rename column to be appropriate
rownames(RNA_vol_df) <- RNA_vol_table$V2[2:RNA_vol_len]
colnames(RNA_vol_df) <- "Volatility"
```

```{r}
#Integrate volatility scores into seurat object
RNA_seurat@meta.data$Volatility <- RNA_vol_df$Volatility
```

```{r}
#plot volatility where a lower score (red) represents a cell more likely to change clusters between algorithms
FeaturePlot(RNA_seurat, features = "Volatility", label = FALSE, cols = c("red", "blue"))
```

```{r}
# Saving the seurat object so it can easily be read in for future analysis
dir.create(file.path(wd, "RNA_seurat_analysis_output", "integrated_seurat"), recursive = TRUE)
saveRDS(RNA_seurat, paste(wd, "RNA_seurat_analysis_output/integrated_seurat", "full_seurat.rds", sep="/"))
```

